### 2019-08-28 20_dataset.tex
redo dataset: smaller table

\section{Dataset}\label{dataset}
The dataset we tested on was a custom selected set of open source 3D models found on the internet which was selected to cover a broad range of different types of application and geometry.
Applications range from prototypes, to fixtures and mechanical end-use parts.
The geometry covers a wide range including thin filaments, smooth surfaces, organic shapes, chamfered shapes, small shapes and large shapes.
The models are described in \cref{dataset_description}.

\begin{table}
\caption{3D models used for validation}\label{dataset_description}
\newcommand{\ssize}{\footnotesize}
\begin{tabular}{l l l}
Model Name & Author \\
\hline
\href{https://www.thingiverse.com/thing:26555}{\ssize AirCasting Air Monitor Casing} & \ssize HabitatMap\\
\href{https://www.thingiverse.com/thing:3629434}{\ssize Air hose splitter} & \ssize frizinko\\
\href{https://www.thingiverse.com/thing:1155772}{\ssize Al Hamra Tower} & \ssize TurnerConstructionCompany\\
\href{https://www.thingiverse.com/thing:1498967}{\ssize canon NP-E3 battery cap} & \ssize kosuyoung\\
\href{https://www.thingiverse.com/thing:3567409}{\ssize David} & \ssize Thunk3D\\
\href{https://www.thingiverse.com/thing:3132621}{\ssize Deck Assembly Tool \comment{, Plank Screwing Tool}} & \ssize PSomeone\\
\href{https://www.thingiverse.com/thing:2920060}{\ssize Ender 3 Cable Chain} & \ssize johnniewhiskey\\
\href{https://www.thingiverse.com/thing:2993875}{\ssize Ergonomic Hacksaw Handle} & \ssize mmOne\\
\href{https://www.thingiverse.com/thing:2513922}{\ssize Gap measurement tool} & \ssize ravm84\\
\href{https://www.thingiverse.com/thing:1673030}{\ssize G-Clamp fully printable} & \ssize johann517\\
\href{https://www.youmagine.com/designs/gyroid}{\ssize Gyroid} & \ssize Tim Kuipers\\
\href{https://www.thingiverse.com/thing:1327093}{\ssize 3D Printable Jet Engine} & \ssize CATIAV5FTW\\
\href{https://www.thingiverse.com/thing:2752165}{\ssize Lawn Mower Throttle Replacement} & \ssize Spammington\\
\href{https://www.thingiverse.com/thing:2854328}{\ssize OpenRC F1 Internal gear box mod} & \ssize intoxikated\\
\href{https://www.thingiverse.com/thing:3592328}{\ssize PCB Test Fixture} & \ssize JMadison\\
\href{https://www.thingiverse.com/thing:1078865}{\ssize Pioneer Radio Holder for Ford Focus} & \ssize Perugino\\
\href{https://www.thingiverse.com/thing:34596}{\ssize Replicator Dual Fan Mount} & \ssize aubenc\\
\href{https://www.thingiverse.com/thing:3754728}{\ssize Atuador versão 2 * actuator version 2} & \ssize Caroline Holanda\\
\href{https://www.thingiverse.com/thing:1595179}{\ssize TE Pocket Operator Hard case} & \ssize Salvation76\\
\href{https://www.thingiverse.com/thing:3682303}{\ssize Screw sizer} & \ssize Pierrolalune63\\
\href{http://homepage.tudelft.nl/z0s1z/projects/2017-bone-infill.html}{\ssize Bone-like optimized infill} & \ssize Jun Wu\\
\href{https://www.thingiverse.com/thing:26244}{\ssize Two-Story Spec House} & \ssize pwc-phil\\

\end{tabular}
\end{table}