### 2019-12-04 0_main.tex
clarify widespread use of concentric toolpaths in abstract (reviewer 3.9)

\documentclass[5p,twocolumn,10pt,times]{elsarticle}
\usepackage{amsmath}
\usepackage{hyperref} % added [draft] to avoid compilation issues that happen if a link is split and appears in two pages
%\modulolinenumbers[5]
\addtolength{\textheight}{8mm}
\addtolength{\textwidth}{4mm}
\addtolength{\voffset}{-10mm}
\addtolength{\hoffset}{-3mm}

\bibliographystyle{elsarticle-num-names}

% ACM template
%
%\documentclass[acmtog,anonymous,timestamp,review]{acmart}
%
%\usepackage{booktabs} % For formal tables
%

% My TK added packages and commands

	% for for using hyperref and elsarticle-num-names together in order to get \citeauthor to work
	\makeatletter
	\providecommand{\doi}[1]{%
	  \begingroup
	    \let\bibinfo\@secondoftwo
	    \urlstyle{rm}%
	    \href{http://dx.doi.org/#1}{%
	      doi:\discretionary{}{}{}%
	      \nolinkurl{#1}%
	    }%
	  \endgroup
	}
	\makeatother

	% have multiline subfigure captions be centered
	\usepackage[labelformat=parens]{subcaption} % subfigures
	\captionsetup[subfigure]{justification=centering}
	\captionsetup{subrefformat=parens} % pure refernce subfigure with parentheses: fig.10a and (b)
	%\renewcommand\thesubfigure{(\alph{subfigure})} % refernce subfigure always with parentheses: fig.10(a) and (b)

	\captionsetup[figure]{labelfont={bf},name={Fig.},labelsep=period} % use `Fig.' for figure subscript instead of `Figure'

	\usepackage[export]{adjustbox} % [right] alignment for includegraphics

	\usepackage{rotating} % turn env for rotating text in figures

	\usepackage{wrapfig} % inline figures

	% tables
	\usepackage{multirow} % multicolumn, multirow
	\usepackage{colortbl} % \cellcolor{<color>}
	\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}   %% centered
	\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}m{#1}}  %% right aligned

	\usepackage[capitalise]{cleveref} % automatically add `Fig.'  etc before a reference.

        \usepackage{ amssymb } % \therefore

	\newcommand{\degree}{^\circ}

	\usepackage[binary-units]{siunitx} % mm and stuff
	\sisetup{per-mode = symbol}
	\DeclareSIUnit\pixel{px}

	\usepackage{units} % \nicefrac{3}{8}

	\DeclareMathOperator*{\argmax}{arg\,max}
	\DeclareMathOperator*{\argmin}{arg\,min}

	\DeclareMathOperator{\abs}{abs} % absolute function

	\usepackage{amsthm} % \begin{proof}
	\newtheorem{lemma}{Lemma}[section]
	\theoremstyle{definition}
	\newtheorem{definition}{Definition}[section]

	\usepackage[inline]{enumitem} % inline enumerate*

	\usepackage[toc,page]{appendix} % appendicces

	\usepackage{pgfplots}
	\usepackage{pgfplotstable} % tikzpicture table plots
	\pgfplotsset{compat=1.15}
	\usetikzlibrary{backgrounds}

	\usepackage[noend]{algpseudocode} % algorithmic
	\usepackage{algorithm} % wrapper for pseudocode to give a caption and label

	\newcommand{\pluseq}{\mathrel{+}=} %pluseq symbol
	\usepackage{upgreek} % \uplambda

	\usepackage{listings} % for listing C++ code instead of pseudocode
	\lstset{ 
      breaklines=true,                 % sets automatic line breaking
      basicstyle=\ttfamily,
      mathescape
    }

    % \usepackage[disable]{todonotes} % notes not showed  
    % \usepackage[draft]{todonotes}   % notes showed
    \usepackage{color,soul} % caps, highlight (\hl)

	\newcommand{\comment}[1]{}

%    \newcommand{\todo}[1]{\hl{#1}}
%    
%	\newcommand{\temp}[1]{\textcolor[rgb]{0, 0, 0.2}{#1}}
%	\newcommand{\tim}[1]{\temp{\todo{[Tim: #1]}}}
%	\newcommand{\jun}[1]{\temp{\todo{[Jun: #1]}}}
%	
%	\newcommand{\old}[1]{\textcolor{gray}{#1}}
	\usepackage[normalem]{ulem}
	\newcommand{\stkout}[1]{\ifmmode\text{\sout{\ensuremath{#1}}}\else\sout{#1}\fi}

	% Revise macro (usage: \revise{old}{new})
	% Version a) First arg red and striked out, second argument green
	\newcommand{\revise}[2]{\textcolor{red}{\stkout{#1}}\textcolor{blue}{#2}}
	% Version b) First arg ignored, second argument green
	%\newcommand{\revise}[2]{\textcolor{blue}{#2}}
	% Version c) First arg ignored, second argument unchanged (for final draft)
	%\newcommand{\revise}[2]{#2}

	\setulcolor{red}

	\usepackage[normalem]{ulem} % squigly underline

	\renewcommand\floatpagefraction{.8}

	\newlength{\figwidth}
	\newlength{\figwidthTwo}
	\newlength{\figwidthTree}
	\newlength{\figheight}
	\newlength{\figheightTwo}
	\newlength{\tempheight}
	\newlength{\tempheightTwo}

	% deal with missing images which are not directly included in the repo
	\iffalse
	\newcommand{\noimage}[1]{%
	  \setlength{\fboxsep}{-\fboxrule}%
	  \fbox{\phantom{\rule{10pt}{10pt}} Missing file: \path{#1} \phantom{\rule{10pt}{10pt}}}% Framed box
	}
	\let\includegraphicsoriginal\includegraphics
	\renewcommand{\includegraphics}[2][width=\textwidth]{\IfFileExists{#2}{\includegraphicsoriginal[#1]{#2}}{\noimage{#2}}}

	\fi
% ENd of TK's added packages and commands

\begin{document}
\baselineskip11pt

\begin{frontmatter}

\title{A framework for adaptive width control of dense contour-parallel toolpaths in \revise{additive manufacturing}{fused deposition modeling}}

%\author{Paper ID: xxx}

\author[um,tud]{Tim Kuipers}
\author[tud]{Eugeni L. Doubrovski}
\author[tud]{Jun Wu\corref{cor1}}
\ead{j.wu-1@tudelft.nl}
\cortext[cor1]{Corresponding author}
\author[cuhk]{Charlie Wang}
% \ead{cwang@mae.cuhk.edu.hk}
\address[um]{Ultimaker, Utrecht, The Netherlands}
\address[tud]{Department of Design Engineering, Delft University of Technology, The Netherlands}
\address[cuhk]{Department of Mechanical and Automation Engineering, The Chinese University of Hong Kong, Hong Kong SAR, China}

\begin{abstract}
3D printing techniques such as Fused Deposition Modeling (FDM) have enabled the fabrication of complex geometry quickly and cheaply. 
\revise{By densely filling consecutive 2D layers with contour-parallel extrusion toolpaths, FDM can produce parts with high stiffness and strength.}
{High stiffness parts are produced by filling the 2D polygons of consecutive layers (partly) with contour-parallel extrusion toolpaths.}
Toolpath with uniform inward offsets from the outline polygons produce over- and underfill regions in the center of the shape, which are especially problematic for thin parts such as casings and microstructures.
Existing approaches for generating toolpaths with adaptive width result in a large variation in widths, which for some hardware systems is difficult to realize accurately, if not beyond their capabilities.
In this paper we present a framework which supports multiple schemes to generate toolpaths with adaptive width, by using a function to decide the number of beads and their widths which are applied from the center outward.
Furthermore, we propose a novel scheme for FDM printing which avoids extreme bead width deviation from the nozzle size, and limits the number of altered toolpaths.
We validate the effectiveness of our framework and this novel scheme on a data set of 300 slices.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
%\begin{CCSXML}
%\end{CCSXML}

%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}

\begin{keyword} 
Adaptive extrusion width, Toolpath generation, Additive manufacturing, Geometrical accuracy, Medial axis transform
\end{keyword}

\end{frontmatter}

%  \temp{%Table of contents just for reviewing purposes
%  \tableofcontents
%  }

\input{1_intro}

\input{2_related_work}

\input{4_method}

\input{5_generalization}

\input{6_validation}

\input{8_discussion}

\input{12_conclusions_future_work}

\section*{References}
\bibliography{99_mybib}

\begin{appendices}

\input{19_edge_discretization}
\input{20_dataset}
\input{25_accuracy}

\end{appendices}
\end{document}