### 2020-01-25 0_main.tex
remove \revise macro everywhere

\documentclass[5p,twocolumn,10pt,times]{elsarticle}
\usepackage{amsmath}
\usepackage{hyperref} % added [draft] to avoid compilation issues that happen if a link is split and appears in two pages
%\modulolinenumbers[5]
\addtolength{\textheight}{8mm}
\addtolength{\textwidth}{4mm}
\addtolength{\voffset}{-10mm}
\addtolength{\hoffset}{-3mm}

\bibliographystyle{elsarticle-num-names}

% ACM template
%
%\documentclass[acmtog,anonymous,timestamp,review]{acmart}
%
%\usepackage{booktabs} % For formal tables
%

% My TK added packages and commands

	\newif\ifcensor
	\newif\ifcolorrevise

	\censorfalse
	\colorrevisetrue

    \usepackage{censor}
	\def\censorcolor{black}
	\let\svcensorrule\censorrule
	\renewcommand\censorrule[1]{%
	\textcolor{\censorcolor}{\svcensorrule{#1}}}

    \newcommand{\censoromit}[1]{}
	\newcommand{\censornotice}{}
	\ifcensor
		\renewcommand{\censornotice}{\\Anonymized 2020 (under revision)}
	\else
		\renewcommand{\censoromit}[1]{#1}
		\StopCensoring
    \fi

	% for for using hyperref and elsarticle-num-names together in order to get \citeauthor to work
	\makeatletter
	\providecommand{\doi}[1]{%
	  \begingroup
	    \let\bibinfo\@secondoftwo
	    \urlstyle{rm}%
	    \href{http://dx.doi.org/#1}{%
	      doi:\discretionary{}{}{}%
	      \nolinkurl{#1}%
	    }%
	  \endgroup
	}
	\makeatother

	% have multiline subfigure captions be centered
	\usepackage[labelformat=parens]{subcaption} % subfigures
	\captionsetup[subfigure]{justification=centering}
	\captionsetup{subrefformat=parens} % pure refernce subfigure with parentheses: fig.10a and (b)
	%\renewcommand\thesubfigure{(\alph{subfigure})} % refernce subfigure always with parentheses: fig.10(a) and (b)

	\captionsetup[figure]{labelfont={bf},name={Fig.},labelsep=period} % use `Fig.' for figure subscript instead of `Figure'

	\usepackage[export]{adjustbox} % [right] alignment for includegraphics

	\usepackage{rotating} % turn env for rotating text in figures

	\usepackage{wrapfig} % inline figures

	% tables
	\usepackage{multirow} % multicolumn, multirow
	\usepackage{colortbl} % \cellcolor{<color>}
	\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}   %% centered
	\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}m{#1}}  %% right aligned

	\usepackage[capitalise]{cleveref} % automatically add `Fig.'  etc before a reference.

        \usepackage{ amssymb } % \therefore

	\newcommand{\degree}{^\circ}

	\usepackage[binary-units]{siunitx} % mm and stuff
	\sisetup{per-mode = symbol}
	\DeclareSIUnit\pixel{px}

	\usepackage{units} % \nicefrac{3}{8}

	\DeclareMathOperator*{\argmax}{arg\,max}
	\DeclareMathOperator*{\argmin}{arg\,min}

	\DeclareMathOperator{\abs}{abs} % absolute function

	\usepackage{amsthm} % \begin{proof}
	\newtheorem{lemma}{Lemma}[section]
	\theoremstyle{definition}
	\newtheorem{definition}{Definition}[section]

	\usepackage[inline]{enumitem} % inline enumerate*

	\usepackage[toc,page]{appendix} % appendicces

	\usepackage{pgfplots}
	\usepackage{pgfplotstable} % tikzpicture table plots
	\pgfplotsset{compat=1.15}
	\usetikzlibrary{backgrounds}

	\usepackage[noend]{algpseudocode} % algorithmic
	\usepackage{algorithm} % wrapper for pseudocode to give a caption and label

	\newcommand{\pluseq}{\mathrel{+}=} %pluseq symbol
	\usepackage{upgreek} % \uplambda

	\usepackage{listings} % for listing C++ code instead of pseudocode
	\lstset{ 
      breaklines=true,                 % sets automatic line breaking
      basicstyle=\ttfamily,
      mathescape
    }

    % \usepackage[disable]{todonotes} % notes not showed  
    % \usepackage[draft]{todonotes}   % notes showed
    \usepackage{color,soul} % caps, highlight (\hl)

	\newcommand{\comment}[1]{}

    \newcommand{\todo}[1]{\hl{#1}}

	\newcommand{\temp}[1]{\textcolor[rgb]{0, 0, 0.2}{#1}}
	\newcommand{\tim}[1]{\temp{\todo{[Tim: #1]}}}
	\newcommand{\jun}[1]{\temp{\todo{[Jun: #1]}}}

	\newcommand{\old}[1]{\textcolor{gray}{#1}}
	\usepackage[normalem]{ulem}
	\newcommand{\stkout}[1]{\ifmmode\text{\sout{\ensuremath{#1}}}\else\sout{#1}\fi}

\iffalse

	% Revise macro (usage: \revise{old}{new})
	\newcommand{\revise}[2]{#2}
	% Version a) First arg red and striked out, second argument green
	%\newcommand{\revise}[2]{\textcolor{red}{\stkout{#1}}\textcolor{blue}{#2}}
	%\newcommand{\revise}[2]{{\color{red}{#1}\color{blue}{#2}}}
	% Version b) First arg ignored, second argument green
	\ifcolorrevise
	\renewcommand{\revise}[2]{{\color{blue}{#2}}}
	\fi
	% Version c) First arg ignored, second argument unchanged (for final draft)
	%\newcommand{\revise}[2]{#2}
	%\newcommand{\revise}[2]{#1}

\fi

	\setulcolor{red}

	\usepackage[normalem]{ulem} % squigly underline

	\renewcommand\floatpagefraction{.8}

	\newlength{\figwidth}
	\newlength{\figwidthTwo}
	\newlength{\figwidthTree}
	\newlength{\figheight}
	\newlength{\figheightTwo}
	\newlength{\tempheight}
	\newlength{\tempheightTwo}

	% deal with missing images which are not directly included in the repo
	\iffalse
	\newcommand{\noimage}[1]{%
	  \setlength{\fboxsep}{-\fboxrule}%
	  \fbox{\phantom{\rule{10pt}{10pt}} Missing file: \path{#1} \phantom{\rule{10pt}{10pt}}}% Framed box
	}
	\let\includegraphicsoriginal\includegraphics
	\renewcommand{\includegraphics}[2][width=\textwidth]{\IfFileExists{#2}{\includegraphicsoriginal[#1]{#2}}{\noimage{#2}}}

	\fi
% ENd of TK's added packages and commands

\begin{document}
\baselineskip11pt

\begin{frontmatter}

\title{A framework for adaptive width control of dense contour-parallel toolpaths in {fused deposition modeling}\censornotice}

%\author{Paper ID: xxx}

\author[um,tud]{\censor{Tim Kuipers}}
\author[tud]{\censor{Eugeni L. Doubrovski}}
\author[tud]{\censor{Jun Wu}\corref{cor1}}
\censoromit{\ead{j.wu-1@tudelft.nl}}
\cortext[cor1]{Corresponding author}
\author[cuhk]{\censor{Charlie Wang}}
% \ead{cwang@mae.cuhk.edu.hk}
\address[um]{\censor{Ultimaker, Utrecht, The Netherlands}}
\address[tud]{\censor{Department of Sustainable Design Engineering, Delft University of Technology, The Netherlands}}
\address[cuhk]{\censor{Department of Mechanical and Automation Engineering, The Chinese University of Hong Kong, Hong Kong SAR, China}}

\begin{abstract}
3D printing techniques such as Fused Deposition Modeling (FDM) have enabled the fabrication of complex geometry quickly and cheaply.

{High stiffness parts are produced by filling the 2D polygons of consecutive layers (partly) with contour-parallel extrusion toolpaths.}
{Toolpaths consisting of \emph{uniform}} inward offsets from the outline polygons produce over- and underfill regions in the center of the shape, which are especially {detrimental to the mechanical performance of} thin parts such as casings and microstructures.
Existing approaches for generating toolpaths with adaptive width result in a large variation in widths, which for some hardware systems is difficult to realize accurately, if not beyond their capabilities.
In this paper we present a framework which supports multiple schemes to generate toolpaths with adaptive width, by using a function to decide the number of beads and their widths which are applied from the center outward.
Furthermore, we propose a novel scheme for FDM printing which {reduces} extreme bead width deviation, {while limiting} the number of altered toolpaths.
We validate the effectiveness of our framework and this novel scheme on a data set of 300 {layer outlines}.
We {statistically }validate the effectiveness of our framework and this novel scheme on a data set of {representative 3D models, }%sgd
{and physically validate it by developing a technique, called \emph{back pressure compensation}, for off-the-shelf FDM systems to effectively realize adaptive width.}
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
%\begin{CCSXML}
%\end{CCSXML}

%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}

\begin{keyword} 
Adaptive extrusion width, Toolpath generation, Additive manufacturing, Geometrical accuracy, Medial axis transform
\end{keyword}

\end{frontmatter}

%  \temp{%Table of contents just for reviewing purposes
%  \tableofcontents
%  }

\input{1_intro}

\input{2_related_work}

\input{4_method}

\input{5_generalization}

\input{6_validation}

\input{8_discussion}

\input{12_conclusions_future_work}

\section*{References}
\interlinepenalty=100000 % prevents pdfendlink ended up accross pages error. see https://tex.stackexchange.com/a/449633/129190
\bibliography{99_mybib}

\begin{appendices}

\input{19_edge_discretization}
\input{20_dataset}
\input{25_accuracy}

\end{appendices}
\end{document}